---
title: "Class05: Data Vis with ggplot"
author: "Tyler Sy (A16651122)"
format: pdf
---

# Graphics systems in R

There are many graphics systems in R for making plots and figures.

We have already played a little with **"base R"** graphics and the `plot()` function.

Today we will start learning about a popular graphics package called `ggplot2()`.

This is an add on package - i.e. we need to install it. I install it (like I install any package) with the `install.packages()` function.

```{r}
plot(cars)
```

Before I can use the functions from a package I have to load up the package from my "library". We use the `library(ggplot2)` command to load it up.

``` {r}
#install.packages('ggplot2')
library(ggplot2)
ggplot(cars)
```

Every ggplot is made up of at least 3 things:
- data (the numbers etc. that will go into your plot)
- aes (how the columns of data map to the plot aesthetics)
- geoms (how the plot actually looks, points, bars, lines, etc.)


```{r}
ggplot(cars) +
  aes(x = speed, y = dist) +
  geom_point()
```

For simple plots, ggplot is more verbose - it takes more code than base R plot.

Add some more layers to our ggplot:

```{r}
ggplot(cars) +
  aes(x = speed, y = dist) +
  geom_point() +
  geom_smooth(method = 'lm', level = 0) +
  labs(title = "Stopping distance of old cars",
       subtitle = "A silly example plot") +
  theme_bw()
```

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
nrow(genes)
ncol(genes)
table(genes$State)
round(table(genes$State) / nrow(genes) * 100, 2)
```

```{r}
ggplot(genes) +
  aes(x = Condition1, y = Condition2) +
  geom_point()
```

```{r}
p <- ggplot(genes) + 
    aes(x=Condition1, y=Condition2, col=State) +
    geom_point() +
    labs(title = "Gene Expression Changes Upon Drug Treatment")
p
p + scale_colour_manual( values=c("blue","gray","red") )

```

```{r}
#install.packages('gapminder')
#install.packages('dplyr')

library(gapminder)
library(dplyr)
gapminder_2007 <- gapminder %>% filter(year==2007)
```

```{r}
ggplot(gapminder_2007) +
  aes(x = gdpPercap, y = lifeExp) +
  geom_point(alpha = 0.5)
```

```{r}
ggplot(gapminder_2007) + 
  geom_point(aes(x = gdpPercap, y = lifeExp,
                 size = pop), alpha=0.5) + 
  scale_size_area(max_size = 10)
```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957)
ggplot(gapminder_1957) +
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent, size = pop),
             alpha = 0.7) + 
  scale_size_area(max_size = 15)

```

```{r}
gapminder_1957 <- gapminder %>% filter(year==1957 | year==2007)

ggplot(gapminder_1957) + 
  geom_point(aes(x = gdpPercap, y = lifeExp, color = continent, size = pop),
             alpha = 0.7) + 
  scale_size_area(max_size = 15) +
  facet_wrap(~year)
```

```{r}
gapminder_top5 <- gapminder %>% 
  filter(year==2007) %>% 
  arrange(desc(pop)) %>% 
  top_n(5, pop)

gapminder_top5
```

```{r}
ggplot(gapminder_top5) + 
  geom_col(aes(x = country, y = pop))
```